---
title: "BABS 508 - Team 007"
Team: Team 007
Authors: Isha Goyal, Islam Shaalan, Priyanshu Najar, Syed Arsalan Raza
Date: "Last modified `r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_depth: 2
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

#Team
Isha Goyal,  Islam Shaalan, Priyanshu Najar, Syed Arsalan Raza

#Introduction
The bla bla bl a is for bla bla
## The Dataset
## Objective
## Literature Review
## Prerequisites

#Setting up the environment and data import

## Loading Libraries
```{r Loading Packages}
library(ggplot2)
library(Hmisc)
library(dplyr)
library(rms)
library(car)
library(visdat)
library(readxl)
library(finalfit)
library(moderndive)
library(corrplot)
```

## Loading Employee Attrition Dataset: `attrition_data`

```{r Loading file and setting env variables}
setwd("/Users/islamshaalan/OneDrive - UBC/Courses/P3/BABS 508/BABS 508 - Group Project")
attrition_data <- read_excel("data/Attrition Sample 7.xlsx")
```

Displaying sample of the dataset

```{r displaying_attr_df}
knitr::kable(head(attrition_data,5),caption="Raw data (first 5 rows)")
#head(attrition_data)
```
Dataset shape is `dim(attrition_data)`

#Understanding the data

## First we have a look at the data to see what is the data type for each feature and what values does it hold
```{r checking the variables}
glimpse(attrition_data)
```
OK! so some of our variables need re-coding

## Data types and recoding
```{r converting_to_categorical}
attrition_data$BusinessTravel <- factor(attrition_data$BusinessTravel)
attrition_data$Department <- factor(attrition_data$Department)
attrition_data$EducationField <- factor(attrition_data$EducationField)
attrition_data$Gender <- factor(attrition_data$Gender)
attrition_data$JobRole <- factor(attrition_data$JobRole)
attrition_data$Attrition = ifelse(attrition_data$Attrition=='No',0,1)
attrition_data$Attrition = as.factor(attrition_data$Attrition)
```

Now that we have the variables clean and tidy

## Describing the variable
```{r describe}
desc_attrition_data <- describe(attrition_data)
html(desc_attrition_data, size=120, scroll = TRUE)
```

```{r check_for_missing_data}

missing_data <- attrition_data[!complete.cases(attrition_data),]
print(paste0('number of missing rows: ', nrow(missing_data)))
#Visualize missing data if the dataset has any missing rows
#attrition_data %>% missing_plot(title='Attrition dataset missing values')
```

Good our data is already clean

Let's check for any duplicate rows
```{r check_duplicates}
dups <- length(attrition_data)  - length(unique(attrition_data))
```
Since we can have multiple employees with the same conditions the important thing is that we don't have duplicates of the same employee. So we are checking if there are any duplicated EmployeeNum. 

Number of duplicates is `dups`

Good we have no duplicate records

We are ready to move on to EDA

#EDA

## Univariate Data

```{r organizing_variables}
cat_attrition_vars = c('BusinessTravel','Department','EducationField','Gender','YearsSinceLastPromotion','YearsWithCurrManager','JobRole','JobLevel','JobSatisfaction','TrainingTimesLastYear', 'WorkLifeBalance','PerformanceRating')
```
### Attrition variable fill

```{r plotting_attrition_distribution}
par(mfrow=c(1,2)) 
ggplot(data = attrition_data, aes(x = Age, color = Attrition)) + geom_histogram(bins = 20, position='fill')
ggplot(data = attrition_data, aes(x = Age, color = Attrition)) + geom_histogram(bins = 20, position='stack')
par(mfrow=c(1,1)) 
```

```{r data_distribution}
#atttrition_data %>% select(BusinessTravel,Gender, EducationField, YearsSinceLastPromotion,JobLevel,JobRole,JobSatisfaction) %>% hist()
attrition_data %>% select(all_of(cat_attrition_vars)) %>% hist()
#hist(attrition_data)
```

```{r }
p <- plot(desc_attrition_data)
p$Categorical
```
!!!!!!!!!!!!!!!!!!!!Describe Data Distribution!!!!!!!!!!!!!!!!!!!!


```{r }
p <- plot(desc_attrition_data)
p$Continuous
```
!!!!!!!!!!!!!!!!!!!!Describe Data Distribution!!!!!!!!!!!!!!!!!!!!


BIVARIATE DATA DISTRIBUTION
```{r cross tab}
xtabs(~Attrition + Age, data = attrition_data)
xtabs(~Attrition + YearsSinceLastPromotion, data = attrition_data)
xtabs(~Attrition + YearsWithCurrManager, data = attrition_data)
xtabs(~Attrition + PercentSalaryHike, data = attrition_data)
xtabs(~Attrition + WorkLifeBalance, data = attrition_data)
xtabs(~Attrition + Gender, data = attrition_data)
xtabs(~Attrition + Education, data = attrition_data)
xtabs(~Attrition + JobRole, data = attrition_data)
xtabs(~Attrition + JobLevel, data = attrition_data)
```

```{r selecting columns}
attrition.sel <- attrition_data %>% select(Age, BusinessTravel, DailyRate, Department,Education, EducationField,	Gender,	
YearsSinceLastPromotion, YearsWithCurrManager, JobLevel,	JobRole,	JobSatisfaction,	MonthlyIncome,
MonthlyRate,	PercentSalaryHike,	PerformanceRating,	TrainingTimesLastYear,	WorkLifeBalance,	Attrition)
```

```{r explore}
attrition.sel$BusinessTravel <- as.numeric(attrition.sel$BusinessTravel)
attrition.sel$Department <- as.numeric(attrition.sel$Department)
attrition.sel$EducationField <- as.numeric(attrition.sel$EducationField)
attrition.sel$Gender <- as.numeric(attrition.sel$Gender)
attrition.sel$JobRole <- as.numeric(attrition.sel$JobRole)
attrition.sel$Education <- as.numeric(attrition.sel$Education)
attrition.sel$JobLevel<- as.numeric(attrition.sel$JobLevel)
attrition.sel$JobSatisfaction<- as.numeric(attrition.sel$JobSatisfaction)
attrition.sel$PerformanceRating<- as.numeric(attrition.sel$PerformanceRating)
attrition.sel$TrainingTimesLastYear<- as.numeric(attrition.sel$TrainingTimesLastYear)
attrition.sel$WorkLifeBalance <- as.numeric(attrition.sel$WorkLifeBalance)



#correlations <- cor(attrition.sel ,method="spearman")
#corrplot(correlations, number.cex = .9, method = "square", 
#         hclust.method = "ward", order = "FPC",
#         type = "full", tl.cex=0.8,tl.col = "black", title = "spearman correlation")
#correlations <- cor(attrition.sel ,method="pearson")
#corrplot(correlations, number.cex = .9, method = "square", 
#         hclust.method = "ward", order = "FPC",
#         type = "full", tl.cex=0.8,tl.col = "black", title = "pearson correlation")

attrition.sel$BusinessTravel <- factor(attrition.sel$BusinessTravel)
attrition.sel$Department <- factor(attrition.sel$Department)
attrition.sel$EducationField <- factor(attrition.sel$EducationField)
attrition.sel$Gender <- factor(attrition.sel$Gender)
attrition.sel$JobRole <- factor(attrition.sel$JobRole)
attrition.sel$JobRole <- factor(attrition.sel$JobRole)
attrition.sel$JobRole <- factor(attrition.sel$JobRole)
attrition.sel$JobRole <- factor(attrition.sel$JobRole)
attrition.sel$Education <- factor(attrition.sel$Education)
attrition.sel$JobLevel<- factor(attrition.sel$JobLevel)
attrition.sel$JobSatisfaction<- factor(attrition.sel$JobSatisfaction)
attrition.sel$PerformanceRating<- factor(attrition.sel$PerformanceRating)
attrition.sel$TrainingTimesLastYear<- factor(attrition.sel$TrainingTimesLastYear)
attrition.sel$WorkLifeBalance <- factor(attrition.sel$WorkLifeBalance)

```


```{r assessing_linearity}
# a <- ggplot(attrition.sel, aes(x = Age, y = Attrition)) +
#   histSpikeg(Attrition ~ Age, lowess = TRUE, data = attrition.sel) +
#   labs(x = "\nAge percentile", y = "Probability(Attrition)\n")
# 
# b <- ggplot(attrition.sel, aes(x = DailyRate, y = Attrition)) +
#   histSpikeg(Attrition ~ DailyRate, lowess = TRUE, data = attrition.sel) +
#   labs(x = "\nDailyRate percentile", y = "Probability(Attrition)\n")
# 
# c <- ggplot(attrition.sel, aes(x = YearsSinceLastPromotion, y = Attrition)) +
#   histSpikeg(Attrition ~ YearsSinceLastPromotion, lowess = TRUE, data = attrition.sel) +
#   labs(x = "\nYearsSinceLastPromotion percentile", y = "Probability(Attrition)\n")
# 
# d <- ggplot(attrition.sel, aes(x = YearsWithCurrManager, y = Attrition)) +
#   histSpikeg(Attrition ~ YearsWithCurrManager, lowess = TRUE, data = attrition.sel) +
#   labs(x = "\nYearsWithCurrManager percentile", y = "Probability(Attrition)\n")
# 
# e <- ggplot(attrition.sel, aes(x = MonthlyIncome, y = Attrition)) +
#   histSpikeg(Attrition ~ MonthlyIncome, lowess = TRUE, data = attrition.sel) +
#   labs(x = "\nMonthlyIncome percentile", y = "Probability(Attrition)\n")
# 
# f <- ggplot(attrition.sel, aes(x = MonthlyRate, y = Attrition)) +
#   histSpikeg(Attrition ~ MonthlyRate, lowess = TRUE, data = attrition.sel) +
#   labs(x = "\nMonthlyRate percentile", y = "Probability(Attrition)\n")
# 
# g <- ggplot(attrition.sel, aes(x = PercentSalaryHike, y = Attrition)) +
#   histSpikeg(Attrition ~ PercentSalaryHike, lowess = TRUE, data = attrition.sel) +
#   labs(x = "\nPercetageSalaryHike percentile", y = "Probability(Attrition)\n")
# 
# a
# b
# c
# d
# e
# f
# g

```
- Transform Data to correct datatype              (DONE)
- Check for duplicates 
- univariate exploration
    - Visualize NAs (final_fit or vis_dat)        (DONE)
    - histogram
    - Check for outliers
    - Check for inaccurate inputs
- bivariate exploration
    - correlations  (vis_cor)
    - histspike data distribution                 (DONE)
    - scatter plots
    - xtabs                                       (DONE)
- Assumption Testing
    -
    -
    -
    - vif test
- BIAS??
- Model
    - Model validation
    - Interpretation
- Visualization


Shouldn't we split the data to train and test?
Confusion Matrix


| Task 	| Start Date 	| End Date 	| Responsible 	| Contributer 	|
|---	|---	|---	|---	|---	|
| Transform Data to correct datatype 	|  	|  	|  	|  	|
| Check for duplicates 	|  	|  	|  	|  	|
| univariate exploration - Visualize NAs (final_fit or vis_dat) - histogram - HIMSC - Check for inaccurate inputs - Check for outliers 	|  	|  	|  	|  	|
| bivariate exploration - correlations  (vis_cor) - scatter plot 	|  	|  	|  	|  	|
| Assumption Testing - - -  -  	|  	|  	|  	|  	|
| BIAS?? 	|  	|  	|  	|  	|
| Model 	|  	|  	|  	|  	|
| Model validation 	|  	|  	|  	|  	|
| Visualization 	|  	|  	|  	|  	|
| Interpretation 	|  	|  	|  	|  	|
|  	|  	|  	|  	|  	|

ggplot(attrition.sel, aes(x = Age, y = Attrition)) +
  histSpikeg(Attrition ~ Age, lowess = TRUE, data = attrition.sel) +
  labs(x = "\nAge percentile", y = "Probability(Attrition)\n")

